---
title: "Challenge 5"
author: "Quinn He"
description: "Introduction to Visualization"
date: "08/22/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_5
  - air_bnb
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

```{r}

abb <- read.csv("_data/AB_NYC_2019.csv")

abb

```

### Briefly describe the data

The data set already looks tidy since each row has only one entry. I can view all the information I will need by looking at one row. I will say the ids are unnecessary for me at this time. The price looks to be on a per night basis, which will be interesting to see when I look at cost of particular boroughs. The calculated_host_listing_count looks like it's the number of other listings a particular user has. I dislike when I see this on airbnb because someone can monopolize certain spots in the city if they have a bunch of rental properties. I may look at this later on to find out if there are companies or individual people listing multiple units.

```{r}

summary(abb)

```

Since this data is relatively tidy, I do not think I will perform any manipulation. I will however make variables to make graphing a little easier.

Below I create a variable to filter out any outliers in the "price" column since the maximum is 10,000. Since the 3rd quartile is 175, I didn't want to go too far from that, but I wanted to include some units that were pretty pricey.

```{r}

price_outli <- filter(abb, price < 600)

```

I just want to check out all the column names here in case I want to refer back to them.

```{r}

colnames(abb)

```

For fun, I wanted to practice group_by again to look at the mean of a few variables per borough.

```{r}

abb %>% group_by(neighbourhood_group) %>% summarise(mean_price = mean(price), mean_reviews = mean(number_of_reviews), mean_availability_per_yr = mean(availability_365))

```

It's kind of as I expected with Manhattan as the most expensive borough to rent out of, followed closely by Brooklyn.

## Univariate Visualizations

```{r}

ggplot(abb, mapping = aes(x = neighbourhood_group)) +
  geom_bar(fill = "grey") +
  labs(title = "# of Unit Listings per Borough", x = "Borough", y = "# of Units")

```

In the above graph, I want to see the amount of Airbnbs per borough. I'm not too surprised by the frequency due to the popularity and tourism in both Manhattan and Brooklyn.

```{r}


ggplot(price_outli, mapping = aes(x = price)) +
  geom_histogram(color = "black") +
  labs(title = "Cost of Airbnbs in NYC", x = "Cost", y = "Airbnb Units")


```

Above I created a histogram to measure out the general cost of Airbnbs in all of NYC. I used my variable "price_outli" to filter out any outliers. It looks like you could get a decent deal on some of the Airbnb units with most of them falling under \$100.

## Bivariate Visualization(s)

Below is a group of scatter plots for each of the five boroughs. I tried to color code them, but could not get the code to work. With these graphs I wanted to look for any correlation between the minimum number of nights that could be rented and its price. It looks like there is no correlation between the two variables.

```{r}

ggplot(price_outli, mapping = aes(x = price, y = minimum_nights, color = "neighbourhood_group")) +
  geom_point() +
  facet_wrap(vars(neighbourhood_group)) +
  labs(title = "Price per Minimum Nights Required", x = "Price per Unit", y = "Minimum Nights")


```
