---
title: "Challenge 2"
author: "Mani Shanker Kamarapu"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - FAOSTAT_cattle_dairy.csv
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in the Data

```{r}
#| label: Reading data
FAOSTAT_cattle_dairy <- read_csv("_data/FAOSTAT_cattle_dairy.csv")
View(FAOSTAT_cattle_dairy)

```

As we read the data, there are many interesting variables and a lot of observations based on yearly analysis of the various cattle products.

## Describe the data

```{r}
#| label : Dimensions of data set
dim(FAOSTAT_cattle_dairy)
```

We can see that there are 36449 rows and 14 columns in the data. We can see from observations from different variables that "Value" variable is pretty much the only real valuable data we are getting from this data set and remaining variables are either unchanged(Domain and Item) or grouping variables(Area, Element, Year and Flag) which can be used to summarize data and find the categories we are interested in and also can be used to find specific information we need and let's try to get to some more useful info using some other functions.

```{r}
#| label: Diff variables of data set
colnames(FAOSTAT_cattle_dairy)
```

These are the variables available in the data set.

```{r}
#| label: Summary
summary(filter(select(FAOSTAT_cattle_dairy, Value)))
```

It is the summary of the Value variable of the data set which is the only real data variable in data set.

```{r}
#| label : Grouping variables
Area_num <- distinct(FAOSTAT_cattle_dairy, Area)
Element_num <- distinct(FAOSTAT_cattle_dairy, Element)
Year_num <- distinct(FAOSTAT_cattle_dairy, Year)
Flag_num <- distinct(FAOSTAT_cattle_dairy, Flag)
Distinct_count <- c(nrow(Area_num), nrow(Element_num), nrow(Year_num), nrow(Flag_num))
Distinct_count
```

We can observe that data set has the data from 1961-2018 on 3 elements at 232 different areas.

I am interested in finding in more information about Flag and how it corresponds to the description they have given and can we group the values by using it.

```{r}
#| label : Flag information
FLag_info <- FAOSTAT_cattle_dairy %>%
  select(Flag,`Flag Description`)
unique(FLag_info)
```

From the above observation we got to know there are 7 types of flags and it is based on data provided and type of data and values they are and how is has been taken.

## Provide Grouped Summary Statistics

#### Area and Element wise analysis

```{r}
AE_analysis <- FAOSTAT_cattle_dairy %>%
  group_by(Area, Element) %>%
  summarise(mean_value=mean(Value, na.rm = TRUE), median_value=median(Value, na.rm = TRUE), sd_value=sd(Value, na.rm = TRUE), min_value=min(Value, na.rm = TRUE), max_value=max(Value, na.rm = TRUE), IQR_value=IQR(Value, na.rm = TRUE))
AE_analysis
```

As per the analysis, in each area the we have three kinds of elements and there are a lot of variations and mean value is not dependent to element itself in any area but it differs in each area and we can get a deeper understanding when we analyse the area and element separately.

#### Year and Element wise analysis

```{r}
YE_analysis <- FAOSTAT_cattle_dairy %>%
  group_by(Year, Element) %>%
  summarise(mean_value=mean(Value, na.rm = TRUE), median_value=median(Value, na.rm = TRUE), sd_value=sd(Value, na.rm = TRUE), min_value=min(Value, na.rm = TRUE), max_value=max(Value, na.rm = TRUE), IQR_value=IQR(Value, na.rm = TRUE))
YE_analysis
```

As per the data, we can say that the elements has been increased by passing year, there are some fluctuations in between but there is an overall increase by the year.

#### Element-wise analysis

```{r}
E_analysis <- FAOSTAT_cattle_dairy %>%
  group_by(Element) %>%
  summarise(mean_value=mean(Value, na.rm = TRUE), median_value=median(Value, na.rm = TRUE), sd_value=sd(Value, na.rm = TRUE), min_value=min(Value, na.rm = TRUE), max_value=max(Value, na.rm = TRUE), IQR_value=IQR(Value, na.rm = TRUE)) %>%
  arrange(desc(mean_value))
E_analysis
```

As per the above observation, the mean value of the production is maximum and yield is the minimum. It shows that the production value is more with the less yield and it leads to an overall profit.

#### Area-wise analysis

```{r}
A_analysis <- FAOSTAT_cattle_dairy %>%
  group_by(Area) %>%
  summarise(mean_value=mean(Value, na.rm = TRUE), median_value=median(Value, na.rm = TRUE), sd_value=sd(Value, na.rm = TRUE), min_value=min(Value, na.rm = TRUE), max_value=max(Value, na.rm = TRUE), IQR_value=IQR(Value, na.rm = TRUE)) %>%
  arrange(desc(mean_value))
A_analysis
```

As per the analysis, Europe has the highest cattle dairy and British Virgin Islands has the minimum, we can say none. 

#### Year-wise analysis

```{r}
Y_analysis <- FAOSTAT_cattle_dairy %>%
  group_by(Year) %>%
  summarise(mean_value=mean(Value, na.rm = TRUE), median_value=median(Value, na.rm = TRUE), sd_value=sd(Value, na.rm = TRUE), min_value=min(Value, na.rm = TRUE), max_value=max(Value, na.rm = TRUE), IQR_value=IQR(Value, na.rm = TRUE)) %>%
  arrange(desc(mean_value))
Y_analysis
```

As per the data, the cattle dairy production has been increased by the passing year, there are some fluctuations in middle but the overall it is increasing, we can see from the data that the increase is nearly 85 percentage from 1961 to 2018.

### Explain and Interpret

I chose Element, Area and Year subgroups from data set. The reason for choosing them is there importance in analysis and also tried to analysis using two variables to know the dependency of values on different variables. So I have done Area and Element wise analysis, Year and Element wise analysis, Element-wise analysis, Area-wise analysis and Year-wise analysis. Conclusion to my analysis would be Europe has the highest of the cattle dairy and the production of the dairy is increasing gradually as by the year. 
