---
title: "Challenge 8"
author: "Emma Rasmussen"
description: "Joining Data"
date: "08/25/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - snl
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(scales)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
if(! require(scales))
  install.packages("scales", repos=
                     "http://cran.us.r-project.org")
```

## Read in data

Reading in the three SNL data sets. I am interested in gender makeup of cast over time, and possibly air time(season fraction) if I have more time.

```{r}
#filter by cast members, and only selecting the columns from this df that I am interested in
snl_actors <- read_csv("_data/snl_actors.csv") %>% 
  filter(type=="cast") %>% 
  select(aid, type, gender)
snl_actors

#selecting columns from snl_casts useful for analysis
snl_casts <- read_csv("_data/snl_casts.csv") %>% 
  select(aid, sid, n_episodes, season_fraction)
snl_casts

#selecting columns date/season info for time variable
snl_seasons <- read_csv("_data/snl_seasons.csv") %>% 
  select(sid, year, n_episodes)
snl_seasons

```

### Briefly describe the data

Data sets include data starting at season 1 of snl, and give information about number of episodes in a season, years of seasons, cast members over the seasons, gender of cast members, episode fraction of cast members etc. There are a couple variables in common, cast member name, and season number. This is where I will join the data sets. I think this data will be useful to look at cast member makeup of seasons over time.

## Tidy Data (as needed)

(see above and after join)

## Join Data

Joining 3 datasets

```{r}
#joining first two datasets
snl1<-snl_actors %>% 
  full_join(snl_casts, by="aid")
snl1

#joining new data set to previous join
snl2<-snl1 %>% 
  full_join(snl_seasons, by="sid")
snl2

```

More Tidying

```{r}
#Reducing columns further to simplify analysis, arranging data in order of season1->46, and removing NA rows
snl2<-snl2 %>% 
  select("gender", "sid") %>% 
  arrange(sid) %>% 
  na.omit(.)
snl2

#grouping data set by season, then generating a count of gender by season
snl2<-snl2 %>% 
  group_by(sid) %>% 
  count(gender)
snl2
```

Analyzing Joined Dataset

```{r}
#Creating a grouped barchart. I use the scales package for this code chunk
ggplot(snl2, aes(fill=gender, y=n, x=sid)) +
  geom_bar(position="dodge", stat="identity", width=.85, space=0.5)+
  labs(x="Season Number", y="Number of Cast Members", fill="Gender", title="Gender Makeup of SNL Cast Members Over Time")+
  scale_y_continuous(breaks= pretty_breaks())+
  scale_fill_manual(values = c("#FF1466", "#9ED6FF", "#B37BFF"))+
  theme_classic()
```

I did not expect a gender disparity in snl casts to persist so long over time. I'm surprised it is not 1:1 or closer to that by now.
