---
title: "Challenge 6 Emma Rasmussen"
author: "Emma Rasmussen"
description: "Visualizing Time and Relationships"
date: "08/23/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_6
  - debt

---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

```{r}
debt<-read_excel("_data/debt_in_trillions.xlsx")
debt
debtOrig<-debt
```

### Briefly describe the data
This data represents debt in trillions over time, broken down by particular categories (ex: mortgage, student loans, credit card debt etc). Assume this is some amalgamation of debt owed by Americans but I could not find the original data source (again I know nothing about economics).

```{r}
print(summarytools::dfSummary(debt,
                        varnumbers = FALSE,
                        plain.ascii  = FALSE, 
                        style        = "grid", 
                        graph.magnif = 0.8, 
                        valid.col    = FALSE),
      method = 'render',
      table.classes = 'table-condensed')
```

## Tidy Data (as needed)

```{r}
#fixing date variable
debtNEW<-debt%>%
  mutate(YearQuarter=parse_date_time(`Year and Quarter`, 
                                     orders="yq"))
debtNEW

#pivoting debt categories
debtNEWPIVOT<-pivot_longer(debtNEW, c("Mortgage", "HE Revolving", "Auto Loan", "Credit Card", "Student Loan", "Other"),
             names_to="DebtCategory", values_to="DebtByCategory")

debtNEWPIVOT

#selecting for 2021 row only. I used tail() because filter was not working for me.
debt2021<-tail(debtNEWPIVOT, n=6)
debt2021

#data frame with 2 variables debt category and debt in trillions for 2021.
select(debt2021, 4, 5)

#creating a new df of my select manually. debt2021 did not work with gg plot so i created a data frame by type in the values which doesn't feel right. Then the graph worked though. I feel like this code is very inefficient
debt2021df <- data.frame(
  name=c("Mortgage","Revolving","Auto Loan","Credit Card","Student Loan", "Other") ,  
  value=c(10.442, 0.322, 1.415, 0.787, 1.570, 0.421)
  )

debt2021df

```

## Time Dependent Visualization
```{r}
ggplot(debtNEW, aes(YearQuarter, Total))+geom_line()+
  labs(x="Date", y="Debt in Trillions", title="Debt 2003-2021")

```
I chose a line graph to help visualize change over time. Easier to visualize increase and deacrease in debt over the date variable
## Visualizing Part-Whole Relationships
```{r}
#Creating bar graph to transform into pie chart
ggplot(debt2021df, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")+
  labs(x="Debt Category", y="Debt in Trillions", title= "2021 Quarter 2 Debt in Trillions")

#trying to create pie chart
ggplot(debt2021df, aes(x=name, y=value)) + 
  geom_bar(stat = "identity")+
  labs(x="Debt Category", y="Debt in Trillions", title= "2021 Quarter 2 Debt in Trillions")+
  coord_polar()
#I don't know what happened. I need help. I was trying to make a simple pie chart
  
```
I chose a pie chart (what was meant to be a pie chart) to help visualize proportion of debt (Bar graph was just precursor to make pie chart with ggplot). I think this paints a clearer picture than a bar graph of how debt falls proportionally by category.