---
title: "Challenge 1"
author: "Mani Shanker Kamarapu"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   **wild_bird_data.xlsx** ⭐⭐⭐
-   StateCounty2012.xlsx ⭐⭐⭐⭐

#### I will be working on the "wild_bird_data" dataset.


```{r}
# Loading `readxl` package
library(readxl)
wild_bird <- read_xlsx("_data/wild_bird_data.xlsx")

# View the dataset
wild_bird

```

## Describe the data

```{r}
#| label: summary
# Use dim() to get dimensions of dataset
dim(wild_bird)
```
There are 147 cases in 2 columns(Reference and Taken from Figure 1 of Nee et al).
Actually the second row has the real column names so we will now make second row as column names and remove the first row.

```{r}
#Rename the column names
colnames(wild_bird) <- wild_bird[1,]
#Removing the first row
wild_bird <- wild_bird[-1,]
#New dimensions of dataset
dim(wild_bird)
#View the dataset
wild_bird
#Summary of dataset
summary(wild_bird)
```
The dataset is in character class so first we need to convert character class to numeric and then get the summary.

```{r}
#Converting datset to numeric
wild_bird$`Wet body weight [g]` <- as.numeric(wild_bird$`Wet body weight [g]`)
wild_bird$`Population size` <- as.numeric(wild_bird$`Population size`)
#Summary of the converted dataset
summary(wild_bird)
```
This is the brief summary of the wild_bird dataset.

