---
title: "Challenge 2 "
author: "Shoshana Buck"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - StateCounty2012
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readxl)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xlsx ⭐

```{r}
StateCounty2012<- read_excel("_data/StateCounty2012.xls") 
StateCounty2012

 


```


## Describe the data

I imported the StateCounty2012.xsl into R and then used the function Subset() to delete columns to tidy the data. Then I "reassigned the data to New_SateCounty2012 and created new column names "state", "county", and "total."

```{r}
#| label: summary
subset(StateCounty2012, select = -c(1,3,5))
New_StateCounty2012=StateCounty2012 
New_StateCounty2012<- StateCounty2012 [-c(1,3,5)] 
rename(New_StateCounty2012, state=1,county=2, total=3 )
new_StateCounty2012<- New_StateCounty2012


```

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

```{r}
new_StateCounty2012 %>% 
  dplyr::group_by(state) %>% 
  summarise(total1=sum(total))
  
```

### Explain and Interpret

Be sure to explain why you choose a specific group. Comment on the interpretation of any interesting differences between groups that you uncover. This section can be integrated with the exploratory data analysis, just be sure it is included.
