---
title: "Challenge 1"
author: "Pavan Datta Abbineni "
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

I've decided to use the `railroad_2012_clean_county.csv` dataset

```{r}
railroadCompleteData<- read_csv("_data/railroad_2012_clean_county.csv")
```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data

```{r head-visualisation}
head(railroadCompleteData)
```
```{r tail-visualisation}
tail(railroadCompleteData)
```
For a dataset to be in `tidy-format` it needs to satisfy the following conditions.  
1)  Each variable has its own column
2)  Each value is in its own cell and
3)  Each observation is located in its own row.  
  
From our visualization of our above dataset we can confidently say that our dataset is already in tidy format.

```{r rows-dimension}
nrow(railroadCompleteData)
```

Our dataset has a total of 2930 rows.

```{r cols-dimension}
ncol(railroadCompleteData)
colnames(railroadCompleteData)
```

We have a total of 3 columns with the names being `state`, `county` and `total_employees`.
```{r}
stateNames = railroadCompleteData$state
countyNames = railroadCompleteData$county
(unique(stateNames))
length(unique(stateNames))
```

We can see that there are 53 unique states data in our dataset.

```{r sorteddata}
tableOfCompleteData =(table(railroadCompleteData$state))
tableOfCompleteData[order(tableOfCompleteData)]
```
We can see that Texas and Georgia are the states with highest employees where as there are quite a few states with fewer than 10 employees.

This data is likely gathered from the official railroad website, as the number of employees currently on payroll is known data to them.
