---
title: "Challenge 1"
author: "Lindsay Jones"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    df-print: paged
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r}
library(readr)
Railroad <- read_csv("_data/railroad_2012_clean_county.csv")
Railroad
```

## Describe the data


Data set contains the number of railroad employees in the United States in 2012, organized by county.
Data was likely gathered reported to labor bureau either by counties or by each railroad station. Each row represents a county. Columns indicate state (or territory), county, and number of employees. There are `r nrow(Railroad)` counties as shown below:

```{r}
#| label: summary
dim(Railroad)
```


The 10 counties with the most railroad employees are:
```{r}
Railroad %>%
  arrange(desc(total_employees)) %>%
  select(state,county,total_employees)%>%
  group_by(total_employees) %>%
  slice(1)%>%
  ungroup()%>%
  arrange(desc(total_employees))%>%
  slice(1:10)
```



State_Railroad_Props illustrates the percentage of railroad workers located in each state or territory.
```{r}
State <- select(Railroad, state)
State_Railroad_Props <- prop.table(table(State))*100
State_Railroad_Props
```

