---
title: "challenge 1"
author: "Shoshana Buck"
desription: "Reading in data and creating a post"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroads_2012_clean_county.csv
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

## Read in the Data

```{r}
railroad <- read_csv("_data/railroad_2012_clean_county.csv")
railroad

```

This data is separated into three columns: state, county, and total employees and 2,930 rows.  

## Describe the data

I imported the data set of railroad_2012_clean_county.csv and renamed it as Railroad. I then used the function colnames() to breakdown the three column names of "state" "county" and "total_employees." From there I used the spec() function to extract the column names. I then used a pipe function in order to filter and select to see the total amount of employees in each state.


```{r}
#| label: summary
colnames(railroad)
spec(railroad)
head(railroad)
railroad %>% 
  group_by(state) %>% 
  summarise(total_employees2=sum(total_employees))
```
