---
title: "Challenge 2 Instructions"
author: "Animesh Sengupta"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.


-   FAOstat\*.csv ⭐⭐⭐

```{r}
#! label: Read Data
FAOstat_livestock <- read_csv("../posts/_data/FAOSTAT_livestock.csv")
```

Add any comments or documentation as needed. More challenging data may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: summary
#Data Dimensions
dim(FAOstat_livestock)
#Data Columns
colnames(FAOstat_livestock)
#Data Preview
head(FAOstat_livestock,3)
```
## Processed Data for analysis and some information about it
```{r}
#! label: Isolate data for analysis
FAOstat_livestock_main <- select(FAOstat_livestock,"Area","Item","Year","Unit","Value")
head(FAOstat_livestock_main,5)
countries_num<-n_distinct(FAOstat_livestock_main$Area)
year_vector<-unique(FAOstat_livestock_main$Year)
unique(FAOstat_livestock_main$Item)
```
The data collected for FAO was between `r year_vector[1]` to `r tail(year_vector,1)` for `r print(countries_num)` Areas in total. 

I am interested in finding the expenditure statistics on livestock for these countries. Lets find out below. 

## Provide Grouped Summary Statistics
Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

```{r}
#| label: Finding Central Tendencies
ct_analysis <- FAOstat_livestock_main %>%
                group_by(Area,Item)%>%
                summarise(mean_val=mean(Value,na.rm = TRUE),
                          median_val=median(Value,na.rm = TRUE),
                          .groups = 'drop')
dim(ct_analysis)
colnames(ct_analysis)
```

##Item wise by analysis
```{r}
#| label: Item wise analysis
ct_item_group <- ct_analysis %>%
  group_by(Item)%>%
  summarise(mean_val=mean(mean_val,na.rm = TRUE))%>%
  arrange(desc(mean_val))

head(ct_item_group,5)
```
As per the data , Cattle generates the most value for any country , with the mean value being `r ct_item_group$mean_val[1]`. We can speculatively attribute this high value to cattle due to its important role in society. For example, a cattle generates value not only be meat consumption but also dairy production. On the other hand, the Mule with lowest mean value of `r tail(ct_item_group$mean_val,1)`, has lesser societal value.

##Area wise analysis
```{r}
#| label: Area wise analysis
ct_area_group <- ct_analysis %>%
  group_by(Area)%>%
  summarise(mean_val=mean(mean_val,na.rm = TRUE))%>%
  arrange(desc(mean_val))

head(ct_area_group,15)
```
Among the countries, Mainland China and India are the biggest mean producer of livestock value. 

##Time wise analysis of largest producer for most valuable livestock
```{r}
#| label: Time series analysis
ct_time_series <- FAOstat_livestock_main %>%
  filter(Item == 'Cattle' & Area == 'China, mainland')%>%
  group_by(Year)%>%
  summarise(mean_val=mean(Value,na.rm = TRUE))%>%
  arrange(Year)

ggplot(data = ct_time_series, aes(x = Year, y = mean_val)) +
  geom_point()
```
As per the time chart plot, the cattle production started to increase substantially during the late 1900s and saw a peak during early 2000. Since then cattle production has a seen a downfall, probably attributing to advent of technology

### Explain and Interpret

For my analysis I chose the subgroup of Area, Item, Time and Value from the FAO Livestock dataset. The reason for choosing such a group of features was because of its high meaningfulness. This would also allow me to conduct Area wise, Item wise and time series Analysis. Conclusion to my analysis would be that Cattle is the MVP while China is the largest producer of livestock.
