---
title: "Challenge 2"
author: "Steve O'Neill"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
df-print: paged
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Introduction

I'm going to try to do hotel_bookings: ⭐⭐⭐⭐

```{r}
hotel_bookings <- read_csv("_data/hotel_bookings.csv")
hotel_bookings
```

Reading in the data is straightforward.

## Describe the data

This dataset is available on [Kaggle](https://www.kaggle.com/datasets/jessemostipak/hotel-booking-demand) and was originally published in the journal [Hospitality Management](https://www.sciencedirect.com/science/article/pii/S2352340918315191).

The data describes two hotels - one 'city' and one 'resort'-style.

Before I begin, I should establish some terminology based on research online:

TA: Travel Agent TO: Tour Operator Distribution channel: HB (meal): Half Board (breakfast + other meal) FB (meal): Full Board (3 meals a day)

## Provide Grouped Summary Statistics

*Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.*

There are just two \[de-identified\] hotels analyzed in this whole dataset:

```{r}
hotel_bookings %>% group_by(hotel) %>% summarise()
```

It is straightforward to give basic descriptive statistics on the Average Daily Rate of both hotels:

```{r}
hotel_bookings %>% select(adr) %>% summary(adr)

```

I can also demonstrate some more interesting stats based on included values like visitor nationality and lead time:

```{r}
by_hotel <- hotel_bookings %>% group_by(hotel)

by_hotel <- by_hotel %>% summarise(
  average.lead.time = mean(lead_time),
  busiest.year = names(which.max(table(arrival_date_year))),
  busiest.month = names(which.max(table(arrival_date_month))),
  most.freq.nationality = names(which.max(table(country))),
  most.infreq.nationality = names(which.min(table(country)))
)

by_hotel
```

As we can see,

-   The city hotel has a longer lead time on average compared to the resort - could this be based on conferences and business travel?
-   Both hotels' busiest year in the dataset was 2016, as well as their busiest month being August \[annually\].
-   The most frequent nationality of guests was that of PRT - Portugal. That makes sense because the hotels are in [Portugal](https://www.sciencedirect.com/science/article/pii/S2352340918315191).
-   The least frequent visitors in the City and Resort hotels hailed from Anguilla and Burundi, respectively.

To demonstrate select(), I pull out a few values:

```{r}
visits <- select(hotel_bookings, hotel, market_segment, children, babies, country, reservation_status, reservation_status_date, arrival_date_month, adr)
visits
```

Using filter(), here are visits that had a higher-than-median amount of babies brought along as compared to other visitors in the same market segment (Direct, Online, Corporate):

```{r}
visits_with_baby <- visits %>% group_by(market_segment) %>% filter(babies > median(babies, na.rm = TRUE))
visits_with_baby
```

So let's compare the general make-up of travelers:

```{r}
table(visits$market_segment)
```

...with those traveling with a baby:

```{r}
table(visits_with_baby$market_segment)
```

Somewhat expectedly, there aren't too many business travelers bringing their kids!
