---
title: "Challenge 4"
author: "Animesh Sengupta"
desription: "More data wrangling: pivoting"
date: "08/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
  - Animesh Sengupta
  - Debt in trillions data
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(stringr)
library(readxl)
library(lubridate)
library(skimr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.csv⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions ⭐⭐⭐⭐⭐

```{r}
library(readxl)
debt_data <- read_excel("../posts/_data/debt_in_trillions.xlsx", .name_repair = "universal")
head(debt_data,10)

```

### Briefly describe the data
The data represents the debt statistics quarterly across different asset class. 

## Tidy Data (as needed)

The data is nearly tidy, we just need to mutate the Year and quarter column and make sure all the numerical values are uniform across the dataset.

```{r}
colnames(debt_data)

debt_data1 <- debt_data%>%
  mutate(
    Date= parse_date_time(Year.and.Quarter,"yq"),
    across(where(is.numeric), round, 4)
  )
head(debt_data1,20)

```

Any additional comments?
So here we converted the 03:Q1 format of date to human readable date using the parse_date_time. Also , we rounded of all the numeric data to 4 decimal places for uniformity.

## Identify variables that need to be mutated

Are there any variables that require mutation to be usable in your analysis stream? For example, are all time variables correctly coded as dates? Are all string variables reduced and cleaned to sensible categories? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

Document your work here.

```{r}

debt_checks<-debt_data1%>%rowwise()%>%
  mutate(Expected_sum=sum(across(.cols=c(Mortgage,HE.Revolving,Auto.Loan,Credit.Card,Student.Loan,Other))),
  Difference=abs(Total-Expected_sum))

head(debt_checks,10)
```

Any additional comments?
As part of Sanity checks, calculation of total debt across column needs to be verified. Across() function was used rowwise to calculate the total debt and compared the absolute value between each other. As per the data, the difference is near to zero hence the total computations are veritable.
