---
title: "Challenge 1 Roy Yoon"
author: "Roy Yoon"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - birds.csv
  - submission 2
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


## Reading Data

```{r reading data file}
library(readr)
birds <- read_csv("_data/birds.csv")

```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## A Quick Look at 'birds'
```{r}
head(birds)
```
## Dimensions

```{r data dimension}
#| label: summary

#understanding the dimensions of data set 'birds'
dim(birds)

```

## Column Names

There are 30977 rows and 14 column in the data set

```{r column names}
#column names in  data set 'birds'

colnames(birds)

```

## Cases when 'birds' data 'Value' is greater then 10000

```{r Value greater than 10000}

#looking at 'birds' data set that has 'Value' column value greater than 10000
more_than_10000 <- filter(birds, Value > 10000)

more_than_10000 

arrange(more_than_10000, Value)

```

The data above shows 'birds' for if the Value column was greater than 10000. The data is arranged by the 'Value' column values.

## Data for Algeria in 'bird'

```{r Algeria Data}
#looking at 'birds' data set that has 'Value' column value greater than 10000 specifically for Algeria 
more_than_10000_ALG <- filter(more_than_10000, Area == "Algeria")

more_than_10000_ALG 
```

The data above examines the data for Algeria by looking at 'birds' data set that haa values greater than 10000.
