---
title: "Challenge 4"
author: "Jerin Jacob"
desription: "More data wrangling: pivoting"
date: "08/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readxl)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


## Read in data

The debt dataset is the category wise debt data in trillions from the first quarter of the year 2003 to the second quarter of 2021.


```{r}
debt <-read_excel("_data/debt_in_trillions.xlsx", skip= 1, col_names = c("Year_Quarter", "Mortgage", "HE_Revolving", "Auto_Loan", "Credit_Card", "Student_Loan", "Other", "Total")) 
debt
  
```
 It has 6 types of debt variables namely Mortgage, Home Equity revolving debt, Auto Loan, Credit Card, Student Loan and a column for all other debts. The total debt got almost doubled over the period from 2003 to 2021.  


## Tidy Data (as needed)

```{r}
debt<-debt%>%
  separate("Year_Quarter", into=c("Year", "Quarter"), sep=":")%>%
  fill(Quarter)
#debt
```


```{r}
debt <- debt %>%
  filter(Quarter == "Q4") %>%
  select(!contains("Quarter"))
#debt
```
The data has been filtered so that we can take the debt details at the year-ending quarter.

```{r}
attach(debt)
par(mfrow=c(3,2))
plot(Year, Mortgage, main="Scatterplot of Year vs. Mortgage")
plot(Year, HE_Revolving, main="Scatterplot of Year vs HE_Revolving")
plot(Year, Auto_Loan, main="Scatterplot of Year vs Auto_Loan")
plot(Year, Credit_Card, main="Scatterplot of Year vs Credit_Card")
plot(Year, Student_Loan, main="Scatterplot of Year vs Student_Loan")
plot(Year, Other, main="Scatterplot of Year vs Other")


```
When the data is being visualized, we get a more clear idea about the trend for each type of debts over the years. Student loan showed a steady increase in the whole period. Mortgage, Auto Loan and Credit card showed an up trend from the year 2003 to 2008 and then had a decline for a couple of years and again started to regain the uptrend. HE Revolving, eventhough doubled in the period till 2008, continuously declined after that. 


## Contribution to the total debt

Each variables' contribution to the total debt is taken as percentage values. 
```{r}
debt_in_percentage <- debt %>%
  mutate(Mortgage = (Mortgage/Total)*100, 
         HE_Rev = (HE_Revolving/Total)*100, 
         AutoLoan = (Auto_Loan/Total)*100, 
         CreditCard = (Credit_Card/Total)*100,
        StudentLoan = (Student_Loan/Total)*100,
        Others = (Other/Total)*100)
debt_in_percentage <- debt_in_percentage %>%
    select(Year, Mortgage, HE_Rev, AutoLoan, CreditCard, StudentLoan, Others)
debt_in_percentage
```


In any given year, Mortgage contributes the highest to the total debt. The contribution of Student Loan is increasing over the years and in 2020 it has reached above 10 % of the total debt. 

