---
title: "Challenge 1"
author: "Ananya Pujary" 
description: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - railroad
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in the Data

```{r}
#| label: reading in the data
railroad <- read_csv("_data/railroad_2012_clean_county.csv")
```

## Describe the data 

I'll be working with the 'railroad_2012_clean_county.csv' dataset.

```{r}
#| label: summary 1

dim(railroad) #describing the 'railroad' dataset's dimensions
```

From this command, we learn that the 'railroad_2012_clean_county.csv' dataset has 3 columns and 2930 rows.

```{r}
#| label: summary 2
colnames(railroad)
head(railroad)
```

The columns in 'railroad' are: 'state' (datatype: character), 'county'(datatype: character), and 'total_employees'(datatype: double class - numeric values with decimal points). These data were probably collected as part of a large-scale survey of the number of railroad employees by county and state in the United States.

```{r}
#| label: summary 3
railroad_arranged <- railroad %>%
  arrange(desc(total_employees)) # arranging data to find the county with the most number of employees
head(railroad_arranged)
```

Cook county in Illinois has the highest number of railroad employees (8207).

```{r}
#| label: summary 4
railroads<- railroad %>%
  group_by(state) %>%  # grouping the data by state
  select(total_employees) %>% # looking only at the 'total_employees' column
  summarize_all(sum, na.rm=TRUE)%>% # adding the number of employees in the counties state-wise
  arrange(desc(total_employees)) # arranging the states from highest to lowest number of employees

head(railroads)
```

Texas has the most railroad employees (19839) and the Armed Forces Pacific has the least (1).
