---
title: "Challenge 1 Instructions"
author: "Meredith Rolfe"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(Hmisc)
library(psych)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xlsx ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
# 
railroad <- read.csv('_data/railroad_2012_clean_county.csv', stringsAsFactors = TRUE, header = TRUE)


```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: summary

# check first 6 rows in the dataset, to get primary understading of the dataset structure
# check first 6 rows in the dataset, to get primary understanding of the dataframe structure

head(railroad)
View(railroad)
dim(railroad) # numbers of rows. columns

#Dataset Description
  # "state" column contains name of different states from United States of America in upper case, abbreviated format
  # "county" column contains names of different counties, for the states in 'state' column, in upper case format
  # "total_employees" column contains state-county wise number of railroad employees.


# struture of railroad dataset

str(railroad)

# Output

  # 'data.frame':	2930 obs. of  3 variables:
  #  $ state          : Factor w/ 53 levels "AE","AK","AL",..: 1 2 2 2 2 2 2 3 3 3 ...
  #  $ county         : Factor w/ 1709 levels "ABBEVILLE","ACADIA",..: 44 33 524 794 963 1415 1417 78 85 95 ...
  #  $ total_employees: int  2 7 2 3 2 1 88 102 143 1 ...

#summary

  # railroad dataset has 2930 rows and 3 columns
  # State column is a factor class; as converted from character class during file read command read.csv() and it has 53 levels.
  # county column has 1709 levels

# Missing Values, check NA value

sum(is.na(railroad))

# check column wise NA values

sum(is.na(railroad$state))
sum(is.na(railroad$county))
sum(is.na(railroad$total_employees))

# Missing Value, check NULL value

sum(is.null(railroad))

# check column wise NULL values

sum(is.null(railroad$state))
sum(is.null(railroad$county))
sum(is.null(railroad$total_employees))

# column, row names of the dataset

colnames(railroad)
row.names(railroad)

# dimension of dataset

dim(railroad) # 2930 observations & 3 columns

# number of rows, number of columns


nrow(railroad)
ncol(railroad)

# Describe & Summary for descriptive analysis
 
 #describe from Hmisc Package

  Hmisc::describe(railroad)

  #describe from psych Pakcage

  psych::describe(railroad)
  
  # summary

  summary(railroad)


```