---
title: "Challenge 6"
author: "Ananya Pujary"
description: "Visualizing Time and Relationships"
date: "08/23/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_6
  - debt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(readxl)
library(skimr)
library(lubridate)
library(summarytools)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

Reading in the 'debt_in_trillions.xslx' dataset.

```{r}
#| label: reading in the data

debt <- read_excel("_data/debt_in_trillions.xlsx")

```

### Briefly describe the data

```{r}
#| label: data description

print(summarytools::dfSummary(debt, varnumbers = FALSE, plain.ascii = FALSE, graph.magnif = 0.50, style = "grid", valid.col = FALSE), 
      method = 'render', table.classes = 'table-condensed')

```
This dataset gives us information on different loan amounts such as student loans, mortgage, and credit card through different years (2003-2021) and quarters. There are 74 rows and 8 columns, of which 1 is character type and 7 are numeric. There aren't any missing values in the columns.

## Tidy Data (as needed)

The 'Year and Quarter' column can be parsed into a date type column.

```{r}
#| label: tidying the data

debt_final <-debt %>%
  mutate(date = parse_date_time(`Year and Quarter`, 
                           orders="yq"))
```

## Time Dependent Visualization

First, I want to look at how total debt varies with time.

```{r}
#| label: time dependent visualization 1

ggplot(data=debt_final) +
  geom_line(aes(x=date, y=Total)) + labs(title = "Total debt increases with time", x = "Time", y = "Total Debt") + theme_minimal()

```

The graph shows a linear increase between 2003 to around 2007, with debt decreasing from 2007 to about 2014, and again increasing through 2021.

Now, looking specifically at how each loan type varies with time:

```{r}
#| label: time dependent visualization 2
debt_long <- debt_final %>%
  pivot_longer(cols = Mortgage:Other, names_to = "loan_type", values_to = "totals") %>%
  select(-Total) %>% mutate(loan_type = as.factor(loan_type))

ggplot(debt_long, aes(x = date,y = log(totals),color = loan_type)) + geom_point() + facet_grid(rows = vars(loan_type)) + labs(title = "Debt Types over the Years", x = "Date", y = "Amount") + theme(strip.text = element_text(size = 6),
         panel.grid.minor = element_blank(), strip.background = element_blank())

```
Mortgage in general is higher than any other loan type.


## Visualizing Part-Whole Relationships

I want to visualize the proportion of each loan type as a whole using a pie chart.

```{r}
#| label: part-whole visualization

debt_long %>% 
  mutate(loan_type = fct_relevel(loan_type, "Mortgage", "HE Revolving",
                            "Auto Loan", "Student Loan",  
                            "Credit Card","Other")) 

debt_grouped <- debt_long %>%
  select(loan_type,totals) %>%
  group_by(loan_type) %>%
  summarize(loantotal = sum(totals)) %>%
   mutate(perc = (loantotal/ sum(loantotal))*100)

ggplot(debt_grouped, aes(x="", y=perc, fill=loan_type)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + labs(title = "Proportion of Loan Types") + theme_void() + scale_fill_brewer(palette="Blues") + geom_text(aes(x = 1.6, label = paste0(round(perc), "%")), position = position_stack(vjust = 0.5))

```
Mortgage takes up the most space proportionally.
