---
title: "Challenge 4"
author: "Pavan Datta Abbineni "
desription: "More data wrangling: pivoting"
date: "08/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

I decided to use the FedFundsRate dataset.

```{r}
FedFundsRateData<-read_csv("_data/FedFundsRate.csv",
                        show_col_types = FALSE)

```

### Briefly describe the data


```{r visualize-dataset}
head(FedFundsRateData)
```
```{r dimensions}
dim(FedFundsRateData)
```

```{r column-names}
colnames(FedFundsRateData)
```

## Tidy Data (as needed)

In tidy data each variable has its own column, every observation its own row and each value its own cell.

The current dataset is not tidy. We need to modify the Year, Month and Day column into a single date column.
Lets create a new column named date which is the addition of the three column names stated above.  



```{r}
FedFundsRateData$Date <- str_c(FedFundsRateData$Year,"-",FedFundsRateData$Month,"-",FedFundsRateData$Day)%>%ymd()%>%as.Date()

```

```{r}
FedFundsRateData<-subset(FedFundsRateData,select=-c(1,2,3))
```

```{r}
FedFundsRateData<-pivot_longer(FedFundsRateData, 1:7, names_to = "Rates", values_to = "Value")
```

```{r}
head(FedFundsRateData)
```


