---
title: "Challenge 5"
author: "Young Soo Choi"
desription: "Introduction to Visualization"
date: "08/22/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_5
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

Read Autralian Marriage tidy data file.

```{r}
ab_nyc <- read_csv("_data/AB_NYC_2019.csv",
                   show_col_types = FALSE)
ab_nyc
```


### Briefly describe the data

This data set looks like a customer review collected by a company. It contains nearly 50,000 people's review and each data distinguished by their id, neighborhood, and etc.

## Univariate Visualizations

First of all, I took a look about their neighbourhood group. I choosed bar graph because it can be recognized easily.

```{r}
count(ab_nyc, neighbourhood_group)
ggplot(ab_nyc, aes(neighbourhood_group)) +
  geom_bar()
```

Most neighbourhood group are Manhattan(21661) and Brooklyn(20104).

Next, I took a closer look at the Manhattan.

```{r}
ab_nyc_man <- ab_nyc%>%
  filter(neighbourhood_group=="Manhattan")

count(ab_nyc_man, neighbourhood)

ggplot(ab_nyc_man, aes(neighbourhood)) +
  geom_bar()
```

32 neighbourhood are in the Manhattan and Harlem's proportion is the largest.

## Bivariate Visualization

Now, I wondered whether availity and amount of review is related. So I made scatter plot regarding availity and reviews per month. I selected review per month to avoid distortion due to subscription period, etc.

```{r}
ggplot(ab_nyc, aes(x=availability_365, y=reviews_per_month)) +
  geom_point()
```

Anything is hard to recognized. It seems like there are no relations between availity and amount of reviews.
