---
title: "Challenge 8"
author: "Nayan Jani"
description: "Joining Data"
date: "08/25/2022"
format:
  html:
    df-print: paged
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_8
  - FAO
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(lubridate)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```




## Read in data


```{r}

birds<-read_csv("_data/birds.csv", col_names = c("Domain_Code", "Domain", "Area_Code", "Area", "Element_Code","Element", "Item_Code","Item", "Year_Code", "Year","Unit", "Value", "Flag", "Flag_Description" ), skip=1)%>%
  select(-c("Year_Code"))
birds

livestock <- read_csv("_data/FAOSTAT_livestock.csv", col_names = c("Domain_Code", "Domain", "Area_Code", "Area", "Element_Code","Element", "Item_Code","Item", "Year_Code", "Year","Unit", "Value", "Flag", "Flag_Description" ), skip=1) %>%
  select(-c("Year_Code"))
livestock

c_groups <- read_csv("_data/FAOSTAT_country_groups.csv",col_names = c("Country_Group_Code","Country_Group","Country_Code", "Country", "M49_Code", "ISO2_Code", "ISO3_Code"), skip=1)
c_groups

c_groups %>%
  count(Country_Group_Code,Country_Code) %>% 
  filter(n > 1)


```


My final filter is what I think the Primary key is for c_groups.

## Join pt1

```{r}



b_sum <-birds %>%
  group_by(Item)%>%
  summarise(avg_stocks = mean(Value, na.rm=TRUE),
            med_stocks = median(Value, na.rm=TRUE),
            n_missing = sum(is.na(Value)))

l_sum <-livestock %>%
  group_by(Item)%>%
  summarise(avg_stocks = mean(Value, na.rm=TRUE),
            med_stocks = median(Value, na.rm=TRUE),
            n_missing = sum(is.na(Value)))

b_sum
l_sum

item_sum <- full_join(b_sum,l_sum)

item_sum

ggplot(item_sum, aes(x=Item, y=avg_stocks))+
  geom_bar(stat="identity") +
  coord_flip() +
  geom_text(aes(label = round(avg_stocks), color = "red"), size=3, hjust=.6) +
  theme(legend.position = "none")
  
  

```


Here I summarized the Birds and Livestock stocks and joined them into one table. I then created a Bar graph showing the average amount of stock for all animals.



```{r}


livestock %>%
  count(Area_Code,Item, Year) %>% 
  filter(n > 1)




livestock_new <- livestock %>%
  mutate(index = row_number()) %>%
  select(index, everything())

c_groups_new <- c_groups %>%
  mutate(index = row_number()) %>%
  select(index, everything())

livestock_new
c_groups_new

```

The First filter represents the possible primary key for livestock. I created a Index for livestock and c_groups so that I could join them together.

## Join pt2

```{r}
livestock %>%
  full_join(birds)

livestock_new %>%
  inner_join(c_groups_new)
  
```



