---
title: "Challenge 2 Will Munson"
author: "Will Munson"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xlsx ⭐
-   FAOstat\*.csv ⭐⭐⭐
-   hotel_bookings ⭐⭐⭐⭐

```{r}
FAOstat <- read_csv("_data/FAOSTAT_livestock.csv")
```

Add any comments or documentation as needed. More challenging data may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

    So, essentially, this data is interpreting the value of livestock around the world. The values are either official or unofficial data, or FAO estimates. 
    
    The data has a total of over 82k slots. Value is the only quantitative variable. Each set of data was recorded between the years of 1961 and 2018. 
```{r}
#| label: summary
summary(FAOstat)
```

## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.

```{r}
FAOstat %>%
  group_by(Year) %>%
  filter(Item == 'Sheep' & `Flag Description` == 'Official data') %>%
  summarize(mean = mean(Value, na.rm = TRUE), sd = sd(Value, na.rm = TRUE))
```

    I added a chart here to get a better understanding of how the mean official data has changed overtime. Just having the numbers here isn't as helpful as having a chart.
```{r}
FAO <- FAOstat %>%
  group_by(Year) %>%
  filter(Item == 'Sheep' & `Flag Description` == 'Official data') %>%
  summarize(mean = mean(Value, na.rm = TRUE), sd = sd(Value, na.rm = TRUE))

plot(x = FAO$Year, y = FAO$mean)
```

### Explain and Interpret

Be sure to explain why you choose a specific group. Comment on the interpretation of any interesting differences between groups that you uncover. This section can be integrated with the exploratory data analysis, just be sure it is included.

    This specific group I chose involved sheep and official data. Since the value changes every year, I chose to take the average value and standard deviation of sheep for each year. What appears to be happening here is the mean is the average value of sheep appears to be relatively stagnant between the years 2000 and 2010. After 2010, the official data shows a major surplus in the value of sheep. 