---
title: "Challenge 8"
author: "Ananya Pujary"
description: "Joining Data"
date: "08/25/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_8
  - faostat
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(summarytools)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

Reading in the faostat datasets.

```{r}
#| label: reading in the data

faostat_egg <- read_csv("_data/FAOSTAT_egg_chicken.csv")
faostat_livestock <- read_csv("_data/FAOSTAT_livestock.csv")
faostat_cattle <- read_csv("_data/FAOSTAT_cattle_dairy.csv")
faostat_country <- read_csv ("_data/FAOSTAT_country_groups.csv")

```

### Briefly describe the data


```{r}
#| label: data description 1

print(summarytools::dfSummary(faostat_egg, varnumbers = FALSE, plain.ascii = FALSE, graph.magnif = 0.50, style = "grid", valid.col = FALSE), 
      method = 'render', table.classes = 'table-condensed')

dim(faostat_egg)
```

The `faostat_egg` dataset has 38170 rows and 13 columns. There are 40 missing values in the 'Values' column and 7548 missing values in 'Flag'.

```{r}

#| label: data description 2

print(summarytools::dfSummary(faostat_livestock, varnumbers = FALSE, plain.ascii = FALSE, graph.magnif = 0.50, style = "grid", valid.col = FALSE), 
      method = 'render', table.classes = 'table-condensed')

dim(faostat_livestock)

```

There are 82116 rows and 13 columns in the `faostat_livestock` dataset. It has missing values too (1301 in 'Value' and 38270 in 'Flag').

```{r}

#| label: data description 3

print(summarytools::dfSummary(faostat_country, varnumbers = FALSE, plain.ascii = FALSE, graph.magnif = 0.50, style = "grid", valid.col = FALSE), 
      method = 'render', table.classes = 'table-condensed')

dim(faostat_country)

```

The `faostat_country` dataset has 1943 rows and 7 columns. It contains information on each country name and unique code, the country groups they belong to, as well as codes for 'M49', 'ISO2', and 'ISO3'. Only the 'ISO2' column has 8 missing values.

## Tidy Data

### Tidying `faostat_country`

`faostat_country` seems tidy enough to work with.

### Tidying `faostat_egg`

```{r}
#| label: tidy data 1

# turning 'Element' and 'Flag Description' into factor type

faostat_egg$Element <- as.factor(faostat_egg$Element)
faostat_egg$'Flag Description' <- as.factor(faostat_egg$'Flag Description')


# deleting 'Year Code'

faostat_egg <- faostat_egg %>%
  select(-'Year Code')

```

### Tidying `faostat_livestock`

```{r}
#| label: tidy data 2

# turning 'Element' and 'Flag Description' into factor type
faostat_livestock$Element <- as.factor(faostat_livestock$Element)
faostat_livestock$'Flag Description' <- as.factor(faostat_livestock$'Flag Description')


# deleting 'Year Code'
faostat_livestock <- faostat_livestock %>%
  select(-'Year Code')

```


### Tidying `faostat_cattle`

```{r}
#| label: tidy data 3

# turning 'Element' and 'Flag Description' into factor type

faostat_cattle$Element <- as.factor(faostat_cattle$Element)
faostat_cattle$'Flag Description' <- as.factor(faostat_cattle$'Flag Description')


# deleting 'Year Code'

faostat_cattle <- faostat_cattle %>%
  select(-'Year Code')

```

## Join Data

Be sure to include a sanity check, and double-check that case count is correct!

Joining `faostat_egg` and `faostat_livestock` by rowbinding:

```{r}
#| label: join data 1

egg_livestock <- bind_rows(faostat_egg,faostat_livestock)

dim(egg_livestock)
```

Both datasets had 13 columns and the combined number of rows in the new dataset is $$38170 + 82116 = 120286$$.

Joining `faostat_cattle` and `faostat_livestock`by rowbinding:

```{r}
#| label: join data 2

cattle_livestock <- bind_rows(faostat_cattle,faostat_livestock)

dim(cattle_livestock)
```

Both datasets had 13 columns and the combined number of rows in the new dataset is $$36449 + 82116 = 118565$$.


```{r}
#| label: join data 3

#figuring out the primary key for 'faostat_country'
faostat_country %>% 
  count('Country Code') %>% 
  filter(n > 1)
faostat_country

egg_livestock %>% 
  count('Area Code') %>% 
  filter(n > 1)

unique(faostat_country$`Country Code`)
unique(egg_livestock$`Area Code`)

```

So 'Country Code' can be used as the primary key for the `faostat_country` dataset to connect to the `egg_livestock` dataset.

```{r}
#| label: join data 4

join_anti <- egg_livestock %>%
  anti_join(faostat_country, by = c("Area Code"="Country Code")) 

dim(join_anti)
unique(join_anti$`Area Code`)
unique(egg_livestock$`Area Code`)

check1 <- join_anti %>% 
  select(`Area Code`,`Area`) %>% 
  group_by(`Area Code`,`Area`) %>% 
  tally()
check1

```

The anti_join() on `egg_livestock` and `faostat_country` reveals that there are 17346 observations in `egg_livestock` that did not find a match in `faostat_country`. `join_anti` has 13 columns, the same as the `egg_livestock` dataset.


## Analyze Data

I'm looking to compare egg production to camel livestock production in Bahrain and Qatar using a line graph with facet wrap.

```{r}
#| label: analyze data 1

egg_livestock_filt <- egg_livestock %>%
  select(Year,Area,Item,Element,Value) %>%
  filter(Area=="Bahrain"|Area=="Qatar") %>%
  filter(Element=="Production"| Item=="Camels")
egg_livestock_filt

```
```{r}
#| label: analyze data 2

egg_livestock_filt %>%
  ggplot(aes(x=Year, y=Value, color=Area)) +
    geom_line() + labs(title = "Egg Production versus Camel Stocks (1961-2018)", x = "Year", y = "Value") + facet_wrap(~ Element) + theme_minimal()

```
Over time, egg production in Qatar exceeded that in Bahrain. Camel livestock production in Qatar evidently far-exceeds stocks in Bahrain too.
