---
title: "Challenge 4"
author: "Young Soo Choi"
desription: "More data wrangling: pivoting"
date: "08/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read Fed Funds Rate

```{r}
ffr<-read_csv("_data/FedFundsRate.csv")
ffr
```

### Briefly describe the data

This data set is regarding various national economic variables from July of 1954 to March of 2017. It has 904 rows and 10 columns. But it also has many missing values and the recorded date is not even constant.

## Tidy Data (as needed)

I Selected data of January 1st and July 1st each year to tidy data. First of all I filtered January 1st data and then filtered July 1st data. And I merged those two data.

```{r}
jan<-filter(ffr, Month==1, Day==1)
jul<-filter(ffr, Month==7, Day==1)
sffr<-rbind(jan,jul)
sffr<-arrange(sffr, Year, Month)
sffr
```
