---
title: "Challenge 2 Instructions"
author: "Young Soo Choi"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readxl)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a data set, and describe the data using both words and any supporting information (e.g., tables, etc)
2)  provide summary statistics for different interesting groups within the data, and interpret those statistics

## Read in the Data

Read in one (or more) of the following data sets, available in the `posts/_data` folder, using the correct R package and command.

-   railroad\*.csv or StateCounty2012.xlsx ⭐
-   FAOstat\*.csv ⭐⭐⭐
-   hotel_bookings ⭐⭐⭐⭐

```{r}
library(readxl)
state2012 <- read_xls("~/R/dacss601/dacss601_august2022/posts/_data/StateCounty2012.xls", skip=3)
state2012
colnames(state2012)
```

This dataset has 2990 rows and 5 columns. And each column name is "STATE", "...2", "COUNTY", "...4", "TOTAL".

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: summary
summary(state2012)
```

This dataset is about the number of workers related to railroad jobs in 2012 I think. And this dataset contains data of 2990 county.


## Provide Grouped Summary Statistics

Conduct some exploratory data analysis, using dplyr commands such as `group_by()`, `select()`, `filter()`, and `summarise()`. Find the central tendency (mean, median, mode) and dispersion (standard deviation, mix/max/quantile) for different subgroups within the data set.
```{r}
MA<-filter(state2012, STATE=="MA")
summary(MA)
CA<-filter(state2012, STATE=="CA")
summary(CA)

```

### Explain and Interpret

MA is the state where I live in and CA is much bigger state I think. MA has 12 county and average total is 286.1 and CA has 55 county and average total is 238.9. So I figure CA has more county than MA but MA's average total is larger than CA's.
