---
title: "Challenge 7"
author: "Young Soo Choi"
desription: "Visualizing Multiple Dimensions"
date: "08/24/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_7
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

Read the data abot organic eggs' price.

```{r}
library(readxl)

eggs <- read_xls("_data/organiceggpoultry.xls",
                 sheet="Data",
                 range="B6:F125",
                 col_names = c("date", "xld", "xlhd", "ld", "lhd"))
eggs
```


## Tidy Data

I tried various methods, but I still couldn't find a suitable method to change the year and month into one variable. So instead, I gave each row an order and tidy the data.

```{r}
eggs<-eggs%>%
  mutate(order=1:120)
eggs
```

## Visualization with Multiple Dimensions

I drew a scatter plot of extra large dozen and large dozen.

```{r}
ggplot(data=eggs)+
  geom_point(mapping=aes(x=xld, y=ld))+
  geom_smooth(mapping=aes(x=xld, y=ld))
```

It looks like they have a strong relation.

The relationship between extra large and its half, large and its half, also seems similar.

```{r}
ggplot(data=eggs)+
  geom_point(mapping=aes(x=xld, y=xlhd))+
  geom_smooth(mapping=aes(x=xld, y=xlhd))
```

```{r}
ggplot(data=eggs)+
  geom_point(mapping=aes(x=ld, y=lhd))+
  geom_smooth(mapping=aes(x=ld, y=lhd))
```

