---
title: "Challenge 5"
author: "Nick Boonstra"
description: "Introduction to Visualization"
date: "08/22/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_5
  - railroads
  - boonstra
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(summarytools)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

```{r read in}

rr_orig<-read_csv("_data/railroad_2012_clean_county.csv")

rr_orig

```

### Briefly describe the data

This data set records railroad employment numbers in the U.S. (and certain overseas locations)

```{r summary}

print(dfSummary(rr_orig, varnumbers = FALSE,
                        plain.ascii  = FALSE, 
                        style        = "grid", 
                        graph.magnif = 0.70, 
                        valid.col    = FALSE),
      method = 'render',
      table.classes = 'table-condensed')

```

One piece of information I was particularly curious about is how average employees per county compared across states:

```{r averages}

rr_orig %>% 
  group_by(state) %>% 
  summarise(mean_emp=mean(total_employees,na.rm=T)) %>% 
  arrange(desc(mean_emp)) %>% 
  slice(1:10)


```

The data show that Delaware has the highest number of average employees per county. This finding becomes even more interesting when investigating how many (or few) counties Delaware has:

```{r counties}

rr_orig %>% 
  filter(state=="DE")

```

Clearly, New Castle county does a lot to offset the mean, especially given that the state of Delaware only has three counties. However, this is not the highest employment in the country:

```{r max_employ}

rr_orig %>% 
  arrange(desc(total_employees)) %>% 
  slice(1:10)



```

Perhaps unsurprisingly, Cook County, IL -- home of a major transit center in Chicago -- employs the most railroad workers of any county in the country. A bit more surprisingly, New Castle County's 1,000+ employees are not actually enough for it to register in the top ten counties!

## Tidy Data (as needed)

These data are already tidy!

## Visualization

Using `ggplot2`, I was able to create a visualization overlaying a density function on top of a histogram of average number of employees per county, when grouped by state.

```{r univariate}

rr_orig %>% 
  group_by(state) %>% 
  summarise(mean_emp=mean(total_employees)) %>% 
  ggplot(aes(x=mean_emp)) +
  geom_histogram(aes(y=..density..),bins=50,alpha=0.5,fill="red") +
  geom_density(fill="blue",alpha=0.2) +
  theme_bw() +
  labs(title="Average Number of Employees per County, by State",
       x="Number of Employees",
       y="Density")
  

```

Because this data set only contains one value, I was not sure how I would create a bivariate visualization.
