---
title: "Challenge 6"
author: "Young Soo Choi"
desription: "Visualizing Time and Relationships"
date: "08/23/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_6
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

Read Fedral Funds Rate data. I changed column names and added date column to make analyse easy.

```{r}
ffr<-read_csv("_data/FedFundsRate.csv",
              show_col_types = FALSE,
              col_names = c("year","month","day","federal_funds_target_rate", "federal_funds_upper_target", "federal_funds_lower_target","effective_federal_funds_rate", "real_gdp", "unemployment_rate","inflation_rate"),
              skip=1)

library(lubridate)

ffr <- mutate(ffr, date = make_datetime(year, month, day))

ffr
colnames(ffr)
```


## Tidy Data

I'd like to check the time series of Real GDP and relation between Real GDP and Unemployment Rate after 2000.
So I filtered after 2000 then selected Real GDP column and Unemployment column. 

```{r}
ffr2<-filter(ffr, year>=2000)
ffr2<-select(ffr2, year, real_gdp, unemployment_rate, date)
ffr2
```


## Time Dependent Visualization

To make time series line graph, I removed the rows that has NA.

```{r}
ffr2<-filter(ffr2, !is.na(real_gdp))
ffr2
```

Then I drawed time line graph of real gdp.

```{r}
ggplot(ffr2, aes(x=date, y=real_gdp)) +
  geom_line()
```

The line is usually between zero and five points during the period, but it hit a low during the financial crisis in the late 2000s.

## Visualizing Part-Whole Relationships

What I concerned is that the relation of real gdp growth and umemployment rate. So I drawed scatter plot of two variable. But according to the scatter plot, it looks like there are no clear relation between those two.

```{r}
ggplot(ffr2, aes(x=real_gdp, y=unemployment_rate))+
  geom_point()
```

