---
title: "Challenge 1"
author: "Young Soo Choi"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

I choose data about birds.

```{r}
library(readr)
birds <- read_csv("_data/birds.csv")
birds
```

This dataset contains over 30000 birds' characters include area, years, etc.

## Describe the data

This data set I read contains data from 1961. I wonder how many birds of that year code is after year 2000 in this data set. 

```{r}
a2000<-filter(birds, Year>=2000)
a2000
count(a2000)
```

And the number of birds with year 2000 after is 10945.
