---
title: "Challenge 3"
author: "Ananya Pujary"
description: "Tidy Data: Pivoting"
date: "08/17/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_3
  - animal_weight
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

library(skimr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in data

I'll be reading in the 'animal_weight' dataset.
```{r}
#| label: reading in the data
animal_weight<-read_csv("_data/animal_weight.csv",
                        show_col_types = FALSE)

```

### Briefly describe the data

```{r}
#| label: data description 1

dim(animal_weight)
colnames(animal_weight) 

```
The data chosen seems to describe the average weights of different animals (dairy and non-dairy cattle, chickens, ducks, etc.) across global regions (Africa, Latin America, Middle East, etc.). It has 9 rows and 17 columns, of which the $n=9$ rows indicate the region name and $k=16$ columns the type of animal. 

```{r}
#| label: data description 2

skim(animal_weight)

```
There are no missing values in this dataset. Overall, dairy cattle seem to have the highest average weight (425.44) and broiler chickens have the lowest (0.9).

I plan to pivot it because it seems that the selected animals are recurring categories in all of the regions. $k-3$ variables will be pivoted and put in a new column.

Hence, the pivoted dataset would have `r 9*16` rows and 3 columns ('IPCC Area','Animal Type', 'Weight').

### Challenge: Describe the final dimensions

```{r}
#| label: describing final dimensions

# existing rows/cases
nrow(animal_weight)

# existing columns/cases
ncol(animal_weight)

#expected rows/cases
nrow(animal_weight) * (ncol(animal_weight)-1)

# expected columns 
1 + 2

```

There are 9 existing rows and 17 existing columns. The expected rows are 144 and expected columns are 3.

### Challenge: Pivot the Chosen Data

```{r}
#| label: pivoting the data

animal_weight_pivoted <- pivot_longer(animal_weight,
                         col = c('Cattle - dairy', 'Cattle - non-dairy', 'Buffaloes', 'Swine - market', 'Swine - breeding', 'Chicken - Broilers', 'Chicken - Layers', 'Ducks', 'Turkeys', 'Sheep', 'Goats', 'Horses', 'Asses', 'Mules', 'Camels', 'Llamas'), names_to = 'Animal Type', values_to = 'Weight')
animal_weight_pivoted

dim(animal_weight_pivoted)
```
The dimensions of the pivoted data, as predicted, are 144 rows and 3 columns. The new case created is 'Animal Type'. Overall, pivoting made the data easier to understand since we can now find the weight of a certain animal from a particular region.
