---
title: "Challenge 1 Sarah Kim"
author: "Sarah Kim"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: tue
categories:
  - challenge_1
  - second attempt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xlsx ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
library(readr)
birds <- read_csv("_data/birds.csv")

```

Add any comments or documentation as needed. More challenging data sets may require additional code chunks and documentation.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: summary

# understanding the dimensions of the birds data set
dim(birds)

# filtering birds data after 2000

BIRDS <- filter(birds, Year > 2000)

# filtering birds data after 2000 in Afghanistan
 
BIRDS_AFG <- filter(BIRDS, Area == "Afghanistan")
BIRDS_AFG

# arranging birds data after 2000 in Afghanistan
arrange(BIRDS_AFG, Value)

```

# General Information about the Data

-   There are 30,977 rows and 14 columns in the birds data set.

-   By filtering the data set so that data after 2000 in Afghanistan, we see that Afghanistan birds data exists between 2001 and 2015.

# Description of the Data

-   By arranging the said data, we see that there are consistent increase in the value from 2014-2016.

-   However, the year and the value do not show a proportional relationship as the highest value is for the year 2005.

-   The two years that did not reach the value of 10,000 are 2001 and 2007.

-   The two years that surpassed the value of 14,000 are 2005 and 2018.
