---
title: "Challenge 3"
author: "Young Soo Choi"
desription: "Tidy Data: Pivoting"
date: "08/17/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_3
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1.  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2.  identify what needs to be done to tidy the current data
3.  anticipate the shape of pivoted data
4.  pivot the data into tidy format using `pivot_longer`

## Read in data

Read the data regarding eggs

```{r}
eggs <- read_csv("_data/eggs_tidy.csv")
eggs
```

### Briefly describe the data

This dataset shows that the price of eggs categorizied by their size from 2004 I think. Because this data has 6 columns, it cannot be easily recognized at once. So I will pivot it. The 4 columns about the types will be pivoted to "types" column.

## Anticipate the End Result

In this 'eggs' dataset 2 of the variables are used to identify a case. So expected rows are 480 and columns are 4.

```{r}
#existing rows/cases
nrow(eggs)

#existing columns/cases
ncol(eggs)

#expected rows/cases
nrow(eggs) * (ncol(eggs)-2)

# expected columns 
2 + 2
```


## Pivot the Data

```{r}
#pivot data

pivot_eggs<-pivot_longer(eggs,col=c(large_half_dozen, large_dozen, extra_large_half_dozen, extra_large_dozen), names_to="types", values_to="price")
pivot_eggs

#change the order of columns
pivot_eggs<-pivot_eggs[c(2,1,3,4)]
pivot_eggs
```

Any additional comments?

It has changed to 480 rows and 4 columns dataset.
