---
title: "Challenge 2"
author: "Ananya Pujary"
description: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - hotel_bookings
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(skimr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in the Data

I'll be working with the 'hotel_bookings.csv' dataset.
```{r}
#| label: reading in the data
hotelbookings <- read_csv("_data/hotel_bookings.csv")

```
## Describe the data

First, we'll generate a broad overview of the data.

```{r}
#| label: summary 1

skim(hotelbookings)

```
This dataset has 32 variables and 119390 cases. There are 14 columns of character type, 1 of date type ('reservation_status_date'), and 18 of numeric type. Only the 'children' column has missing values (4).

```{r}
#| label: summary 2

distinct(hotelbookings,arrival_date_year) 
distinct(hotelbookings,is_canceled)

```

From the variable names, this dataset seems to be logging the data of a hotel's bookings for the years 2015-2017, such as their customers' arrival date, reservation status, and the number of adults/children/babies checking in. The variable 'is_canceled' has binary values and indicates whether a booking has been canceled (0) or not (1).

```{r}
#| label: summary 3

distinct(hotelbookings,reservation_status)

distinct(hotelbookings,hotel) 

hotelbookings %>%
  group_by(hotel) %>%
  tally()
```

The 'reservation_status' variable has three values: 'Check-Out', 'Canceled', and 'No-Show'. There are two types of hotels from which data are collected: 'Resort Hotel' and 'City Hotel'. 79330 of the rows contain data related to the City Hotel, while 40060 rows are related to the Resort Hotel.

## Provide Grouped Summary Statistics

The following command returns central tendency and dispersion values for the numeric values in the dataset.

```{r}
#| label: grouped summary 1
#| 
summary(hotelbookings)
```

```{r}
#| label: grouped summary 2

# calculating the number of cancellations and number of repeat guests grouped by hotel
hotel_stats<-hotelbookings %>%
  select(hotel,is_canceled,is_repeated_guest) %>%
  group_by(hotel)%>%
  summarize(cancellations = sum(is_canceled),guests_repeated = sum(is_repeated_guest))

hotelbookings$room_assignment = (ifelse(hotelbookings$reserved_room_type==hotelbookings$assigned_room_type, 'Same', 'Different'))

rooms <- hotelbookings %>%
  select(hotel,room_assignment) %>%
  group_by(hotel,room_assignment) %>%
  tally()
```

City Hotel saw 33102 cancellations and Resort Hotel had 11122 cancellations during this time period. City Hotel also had more repeat guests (2032) than the Resort Hotel (1778). Also, I created a new column in the dataset called 'room_assignment' that indicates whether guests got their preferred room assignment ('Same') or not ('Different'). City Hotel guests were less likely to be assigned a room not of their preference (7192) than Resort Hotel guests (7725).


```{r}
#| label: grouped summary 3

hotel_reservations <- hotelbookings %>%
  select(hotel,reservation_status)%>%
  group_by(hotel, reservation_status)%>%
  tally()
```

Then, I compared reservation statuses ('Canceled, 'Check-Out','No Show') across hotels. For both, the number of guests checked out were more than those that cancelled or didn't show up.

```{r}
#| label: grouped summary 4

hotelbookings %>% 
  select(hotel,arrival_date_year,arrival_date_month) %>% 
  group_by(hotel) %>% 
  count(arrival_date_year,arrival_date_month) %>%
  slice(which.max(n))

```

I also wanted to look at which month-year was the busiest in terms of bookings for both hotels. The busiest month was May 2017 for City Hotel (4556), and October 2016 for Resort Hotel (1984). 

### Explain and Interpret

I chose certain variables that would objectively indicate which hotel (City Hotel or Resort Hotel) was more successful, such as the number of cancellations and reservation status. However, City Hotel had more entries in the dataset than Resort Hotel, which could skew the results. I also wanted to compare which hotel assigned most customers their preferred room type (reserved_room_type versus assigned_room_type) which could be contributing to customer satisfaction, and found that City Hotel was better in this aspect. It would also be interesting to look at both hotels' statistics through the years (2015-2017) to check whether their businesses have improved or declined. 
