---
title: "Challenge 4 Instructions"
author: "Shoshana Buck"
description: "More data wrangling: pivoting"
date: "08/18/2022"
format:
  html:
    df-print: paged
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
    css: "styles.css"
categories:
  - challenge_4
  - FedFundRate
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)
library(lubridate)
library(summarytools)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


## Read in data

```{r}
fedfundrate <- read_csv("_data/FedFundsRate.csv",
                        show_col_types = FALSE, col_names = c("Year", "Month", "Day", "federal_funds_target_rate", "federal_funds_upper_target", "federal_funds_lower_target", "effective_federal_funds_rate", "real_GDP_percent_change", "unemployment_rate", "inflation_rate"),
                        skip=1
)
fedfundrate

```

### Briefly describe the data
This data has been collected monthly since July 1st of 1954 and ended on march 16, 2017. There is a year-month-day combination that looks at 4 different Federal Funds Rates, while the other variables are indicators that effect the federal fund rate (inflation, unemployment and GDP changes).

## Tidy Data (as needed)
```{r}
colnames(fedfundrate)

print(summarytools::dfSummary(fedfundrate,
                        varnumbers = FALSE,
                        plain.ascii  = FALSE, 
                        style        = "grid", 
                        graph.magnif = 0.70, 
                        valid.col    = FALSE),
      method = 'render',
      table.classes = 'table-condensed')

```

## Description of the variables
I think that the summary tools function is really great because it captures the stats/values within each variable. Additionally it displays a graph of the data from each observation. One thing that I think is interesting to point out is that 

## Identify variables that need to be mutated

```{r}
ffr_long <- fedfundrate %>% 
  pivot_longer(cols = c( "federal_funds_target_rate", "federal_funds_upper_target", "federal_funds_lower_target", "effective_federal_funds_rate", "real_GDP_percent_change", "unemployment_rate", "inflation_rate"),
               names_to = "ffr_type",
               values_to = "value",
               values_drop_na = TRUE)

ffr_long

```
I think that using the function pivot_longer() is a great way to tidy the data and be able to get a better understanding/read of the data.  
