---
title: "Challenge 3 Instructions"
author: "Roy Yoon"
desription: "Tidy Data: Pivoting"
date: "08/17/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_3
  - eggs_tidy.csv
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readxl)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1.  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2.  identify what needs to be done to tidy the current data
3.  anticipate the shape of pivoted data
4.  pivot the data into tidy format using `pivot_longer`

## Read in data

```{r}
eggs<-read_csv("_data/eggs_tidy.csv")
```

### About eggs_tidy.csv data set

```{r}
eggs
colnames(eggs)
dim(eggs)
```

The eggs data set has 120 rows and 6 columns.

The data examines the monthly price of different egg sizes and amounts from 2004 to 2013.

## Anticipate the End Result


### Example: find current and future data dimensions

Lets see if this works with a simple example.

```{r}
#| tbl-cap: Example

df<-eggs

df

#existing rows/cases
nrow(df)

#existing columns/cases
ncol(df)

#expected rows/cases
nrow(df) * (ncol(df)-2)

# expected columns 
5
```

## eggs data set pivot longer (month, year, size, amount, price)
```{r}

eggs_tidy_longer<- eggs%>%
  pivot_longer(cols=contains("large"),
               names_to = c("size", "amount"),
               names_sep="_",
               values_to = "price"
  )
eggs_tidy_longer
```

New variable the size and amount are now separated and independent variables of each other and price is also aviable as its own variable


## Describing Pivoted Data

```{r}
dim(eggs_tidy_longer)

colnames(eggs_tidy_longer)

```
The eggs_tidy_longer data set contains 480 rows and 5 columns.

The columns distinguish month, year, size, amount, and price as variable.

The variables organized in this manner allows us to better understand and analyze how the variable are indepent from each other when looking at, size, amount, and price.



