---
title: "Challenge 1_Jyoti Rani"
author: "Jyoti Rani"
desription: "Reading in data and creating a post"
date: "08/15/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
---

```{r}

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xlsx ⭐⭐⭐⭐

I will be working on the “wild_bird_data” dataset.

```{r}
library(readxl)
wild_bird_data <- read_xlsx("/Users/jyotirani/Documents/r_subdirectory/wild_bird_data.xlsx")

# View the dataset
wild_bird_data
```
## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
# Use dim() to get dimensions of dataset
dim(wild_bird_data)
```
There are 147 cases in 2 columns(Reference and Taken from Figure 1 of Nee et al). Actually the second row has the real column names so we will now make second row as column names and remove the first row.

```{r}
#Rename the column names
colnames(wild_bird_data) <- wild_bird_data[1,]
#Removing the first row
wild_bird_data <- wild_bird_data[-1,]
#New dimensions of dataset
dim(wild_bird_data)
```

```{r}
#View the dataset
wild_bird_data
```
```{r}
#Summary of dataset
summary(wild_bird_data)
```

```{r}
#Converting datset to numeric
wild_bird_data$`Wet body weight [g]` <- as.numeric(wild_bird_data$`Wet body weight [g]`)
wild_bird_data$`Population size` <- as.numeric(wild_bird_data$`Population size`)
#Summary of the converted dataset
summary(wild_bird_data)
```

Brief summary of the wild_bird dataset.