---
title: "Challenge 1"
author: "Adithya Parupudi"
desription: "Reading in data and creating a post"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_1
  - tidyverse
  - birds.csv
  - hw2
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to

1)  read in a dataset, and

2)  describe the dataset using both words and any supporting information (e.g., tables, etc)

## Read in the Data

Read in one (or more) of the following data sets, using the correct R package and command.

-   railroad_2012_clean_county.csv ⭐
-   birds.csv ⭐⭐
-   FAOstat\*.csv ⭐⭐
-   wild_bird_data.xlsx ⭐⭐⭐
-   StateCounty2012.xlsx ⭐⭐⭐⭐

Find the `_data` folder, located inside the `posts` folder. Then you can read in the data, using either one of the `readr` standard tidy read commands, or a specialized package such as `readxl`.

```{r}
birds_data <- read_csv("_data/birds.csv",show_col_types = FALSE)
#spec(birds_data) -> full column specification

```

After importing the csv file, and I notice that out of the 14 columns, 8 of them are of character type and 6 columns are double. Total rows -\> 30977!

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
names(birds_data)

```

Column names at a glance

```{r}
head(birds_data)
```

```{r}
#| label: summary
str(birds_data)
```

We get to see a get a high level view of the column names and its entries.

```{r}
hist(birds_data$`Item Code`)
hist(birds_data$`Area Code`)
```

Using the histogram functions, observed that the frequency for item code and area codes respectively.
