---
title: "Challenge 8 Will Munson"
author: "Will Munson"
description: "Joining Data"
date: "08/25/2022"
format:
  html:
    toc: true
    code-copy: true
    code-tools: true
categories:
  - challenge_8
  - railroads
  - snl
  - faostat
  - debt
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in multiple data sets, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  mutate variables as needed (including sanity checks)
4)  join two or more data sets and analyze some aspect of the joined data

(be sure to only include the category tags for the data you use!)

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

  - faostat ⭐⭐
  - railroads  ⭐⭐⭐
  - fed_rate ⭐⭐⭐
  - debt ⭐⭐⭐
  - us_hh ⭐⭐⭐⭐
  - snl ⭐⭐⭐⭐⭐

```{r}
faostat_country <- read_csv("_data/FAOSTAT_country_groups.csv", show_col_types = FALSE)
faostat_livestock <- read_csv("_data/FAOSTAT_livestock.csv", show_col_types = FALSE)
```

### Briefly describe the data

So, the datasets I'll be using are basically for livestock, and I'll also be adding the country codes to this data to get a better understanding of how these values work. 

## Tidy Data (as needed)

Is your data already tidy, or is there work to be done? Be sure to anticipate your end result to provide a sanity check, and document your work here.

The data, surprisingly enough, is tidy! 

```{r}


```

Are there any variables that require mutation to be usable in your analysis stream? For example, do you need to calculate new values in order to graph them? Can string values be represented numerically? Do you need to turn any variables into factors and reorder for ease of graphics and visualization?

Document your work here.

As far as I'm concerned, I don't see any problems, other than the fact that I might not end up using every variable in this dataset. The only thing that concerns me is the fact that the Country Groups dataset contains the same country in multiple categories, and I'm a bit concerned that this would complicate the process. 
```{r}

```

## Join Data

Be sure to include a sanity check, and double-check that case count is correct!

```{r}
FAOstat <- merge(faostat_livestock, faostat_country, by.x = "Area Code", by.y = "Country Code")

FAOstat
```

The dataset is... really big, but I think it works for my purposes. 

```{r}
FAOstat %>%
  filter(`Flag Description` == 'Official data' & `Country Group Code` == 5305)
```

