---
title: "Challenge 2 Instructions"
author: "Roy Yoon"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
  - railroad
  - question
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
#library(readr)
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Read in the Dataset "railroad_2012_clean_county.csv"


```{r initial inspection}
railroad <- read_csv("_data/railroad_2012_clean_county.csv")

railroad

dim(railroad)

colnames(railroad)
```

There are three variable names: 'state', 'county', and 'total_employees'.

## Describe the data

Using a combination of words and results of R commands, can you provide a high level description of the data? Describe as efficiently as possible where/how the data was (likely) gathered, indicate the cases and variables (both the interpretation and any details you deem useful to the reader to fully understand your chosen data).

```{r}
#| label: summary

colnames(railroad)

dim(railroad)

railroad 
```

## Summary Statistics

First I tried to attempt at making the data grouped by states with a single total employee count(regardless of the counties)
```{r}
state_grouped_railroad <- railroad %>%
    select(state, total_employees) %>%
    group_by(state) %>%
    tally(total_employees)

state_grouped_railroad <-rename(state_grouped_railroad, total_employees = n)

state_grouped_railroad

```

### Explain and Interpret

Work is still in progress.

Thought process:

* examine which state has the most/least 'total_employees'

* identify which county has the most/least 'total_employees' in the state with the most/least 'total_employees'  

* look at overall, which county has the most/least 'total_employees', and how does that compare to the state values

* examine the average, min, max across states/counties
