---
title: "hw 2"
author: "Lai Wei"
desription: "Howmework 2"
date: "08/20/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - hw2
  - Lai Wei
  - dataset
  - ggplot2
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Homework 2: Reading in Data
-Read in a dataset from the _data folder in the course blog repository, or choose your own data. If you decide to use one of the datasets we have provided, please use a challenging dataset - check with us if you are not sure. 
-Clean the data as needed using dplyr and related tidyverse packages.
-Provide a narrative about the data set (look it up if you aren't sure what you have got) and the variables in your dataset, including what type of data each variable is. The goal of this step is to communicate in a visually appealing way to non-experts - not to replicate r-code.
-Identify potential research questions that your dataset can help answer


```{r}
#Import Wild_bird data from pasts
library(readxl)
wild_bird_data <- read_excel("_data/wild_bird_data.xlsx",
                             skip = 2,
                             col_names = c("Body Weight", "Population Size"))
wild_bird_data
```

## describe the data 
```{r}
#Show the colname of Wild_bird_data
colnames(wild_bird_data)
#Get the dimensions of electric
dim(wild_bird_data)
```
## Select the Birds Weight which is greater than 20
Using filter() function to get cases.
```{r}
#Filter the rows in Weight that are greater than 25
filter(wild_bird_data, 'Body Weight' > 25.0)

```
## Arrange Data 
Listing Population values in descending order from highest to lowest. 
```{r}
#Summary of Population Size variables
summary(population)

arrange(wild_bird_data, desc('Population Size'))
  
```

## Homework Goal

Re-arrange data in order, and get some basic information about wild_bird. 
:::