---
title: "Challenge 2"
author: "Saaradhaa M"
description: "Data Wrangling"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
    df-print: paged 
categories:
  - challenge_2
  - hotel_bookings.csv
  - tidyverse
  - readr
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(readr)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Reading in the data

```{r}
#| label: read in
# Read in and view the dataset.
hotel <- read.csv("_data/hotel_bookings.csv")
hotel
```

## Description of data

```{r}
#| label: description
# Get rows and columns.
dim(hotel)
# Find which columns have missing data.
which(colSums(is.na(hotel))>0)
```

In the hotel bookings dataset, there are 119390 cases and 32 columns. Only the **children** column has missing data (*N* = 11). Interesting columns include assigned room type, previous cancellations, days in waiting list and is_canceled. There are also columns for **country** and **hotel**, indicating that the data was likely gathered by surveying different hotels around the world.

## Grouped summary statistics #1

I first want to examine the relationship between number of days in the waiting list and whether the booking was cancelled.

```{r}
#| label: Question 1 Summary Statistics
# Check if is_canceled is binary.
apply(hotel,2,function(x) { all(x %in% 0:1) })

# Check mean and median for days in waiting list.
summarise(hotel, diwl_mean = mean(days_in_waiting_list), diwl_median = median(days_in_waiting_list))

# Check mean of is_canceled, grouped by days in waiting list.
hotel %>%
  group_by(days_in_waiting_list) %>%
  select(`is_canceled`) %>%
summarise(is_canceled_mean = mean(is_canceled))
```

Common sense tells me that those who wait longer are more likely to cancel their reservations, but I want to check if this can actually be observed in our data. The code chunk above demonstrates that is_canceled is a binary variable (with values 0 and 1). The tables generated also show that on average, people spend about 2 days on the waiting list. However, some people were on the waiting list for over a year!

The mean cancellation rate is 1 for those who waited 391 days (understandable), and 0.36 for those who didn't wait at all.

## Grouped summary statistics #1

I also want to examine the relationship between assigned room type and previous cancellations.

```{r}
# Check median previous cancellations, grouped by assigned room type.
hotel %>%
  group_by(assigned_room_type) %>%
  select(previous_cancellations) %>%
summarise(previous_cancellations_median = median(previous_cancellations))
```

Most people assigned types A to K and type P were not likely to have previously cancelled their bookings. However, most people assigned type L were like to have previously cancelled their bookings on one occasion - if we can find the data source, it would be interesting to uncover how type L differs from the other room types (were these people given smaller rooms?).
