---
title: "Challenge 4 Instructions"
author: "Yakub Rabiutheen"
desription: "More data wrangling: pivoting"
date: "08/18/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_4
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Challenge Overview

Today's challenge is to:

1)  read in a data set, and describe the data set using both words and any supporting information (e.g., tables, etc)
2)  tidy data (as needed, including sanity checks)
3)  identify variables that need to be mutated
4)  mutate variables and sanity check all mutations

## Read in data

Read in one (or more) of the following datasets, using the correct R package and command.

-   abc_poll.csv ⭐
-   poultry_tidy.csv⭐⭐
-   FedFundsRate.csv⭐⭐⭐
-   hotel_bookings.csv⭐⭐⭐⭐
-   debt_in_trillions ⭐⭐⭐⭐⭐


Column Names

```{r}
colnames(abc_poll_2021)
```

### Briefly describe the data

## Tidy Data (as needed)

This Data is already tidy, perhaps I will compare data about education and Political Identification.

It is also interesting to note that there seems to be a high frequency of working full-time and having a Bachelors degree.

```{r}
table(abc_poll_2021$ppeducat,abc_poll_2021$ppemploy)
```

Cleaned Up Party id similar to how Meredith did it. Removed all Skipped answers and Removed the An.

```{r}
abc_poll_2021<-abc_poll_2011%>%
  mutate(partyid = str_remove(QPID, "A[n]* "),
         partyid = na_if(partyid, "Skipped"))%>%
  select(-QPID)
table(abc_poll_2021$QID)
```




Table below demonstrates the general trend that Democrats are more likely to have a Bachelor's Degree or Higher

```{r}
table(abc_poll_2021$partyid,abc_poll_2021$ppeducat)
```


```

