---
title: "HW 2 Instructions"
author: "Yakub Rabiutheen"
desription: "Data wrangling: using group() and summarise()"
date: "08/16/2022"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - challenge_2
---

```{r}
#| label: setup
#| warning: false
#| message: false

library(tidyverse)
library(summarytools)

knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```



Loaded the Fed Funds Dataset

```{r}
library(readr)
fedfunds <- read_csv("_data/FedFundsRate.csv")
```

Add any comments or documentation as needed. More challenging data may require additional code chunks and documentation.

## Describe the data

Doing a Head of the Dataset to get a view of what the Data Looks like.


```{r}
head(fedfunds)
```

Getting the column names
```{r}
colnames(fedfunds)
```


Renaming columns as I had difficulty invoking the column name in code.
```{r}
names(fedfunds)[names(fedfunds) == "Real GDP (Percent Change)"] <- "Real GDP"
colnames(fedfunds)
```





Summary of the Datset and found a lot of NAs to clean up.
```{r}
summary(fedfunds)
```

Dropping all NAs
```{r}
year_metrics <- fedfunds %>%
  drop_na(Year,`Real GDP`,`Inflation Rate`,`Unemployment Rate`,`Effective Federal Funds Rate`)
year_metrics
```


```{r}
summary(year_metrics)
```


I tried to use a line graph but it looks messy.
```{r}
library(ggplot2)
ggplot(data=year_metrics, aes(x=`Year`, y=`Real GDP`))+
  geom_line()
```




This Bar Graph came out much cleaner looking.
```{r}
perctgdp<-ggplot(year_metrics, aes(x =`Year`, y =`Real GDP`,fill=`Real GDP`)) + geom_bar(stat = "identity")
perctgdp
```

I added limits and breaks to see the most recent year 2017.
```{r}
perctgdp<-perctgdp+scale_x_continuous(limits=c(1954,2017),breaks=seq(1954,2017,5))
```



Relabeled the Y axis to make to clear this not that actual GDP rate but the Percentage change of the GDP    
```{r}
print(perctgdp + labs(
  title = "GDP Percentage Change by Year",
  y = "GDP Percentage Change", x = "Year"
))
```


```{r}
ggplot(year_metrics) +
  aes(x = `Unemployment Rate`, y = `Inflation Rate`) +
  geom_point(colour = "#0c4c8a") +
  theme_minimal()
```


I did a Correlation test of Unemployment rate and the Inflation rate. It got only a moderate Correlation.
```{r}
cor(year_metrics$`Unemployment Rate`,year_metrics$`Inflation Rate`)
```

Trying to Graph both the Unemployment Rate and Inflation Rate by Year
```{r}
df <- year_metrics %>%
  select(Year, `Unemployment Rate`, `Inflation Rate`) %>%
  gather(key = "variable", value = "value", -Year)
```



Graph that shows inflation rate vs Unemployment Rate by Year

As shown here, there can be periods of High Inflation and low unemployment which is why there isn't a strong correlation.
```{r}
ggplot(df, aes(x =`Year`, y = value)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
  scale_color_manual(values = c("darkred", "steelblue"))+labs(
  title = "Inflation and UnEmployment Rates by Year",
  y = "Variable Rates", x = "Year")
```

